---
title: "mlr3extralearners"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mlr3extralearners)
library(magrittr)
```

# What is mlr3extralearners?

`mlr3extralearners` contains all learners from mlr3 that are not in `mlr3learners` or the
core packages. `mlr3extralearners` contains helper functions to find where all these learners live
and to install required packages to run these learners.

```{r}
list_mlr3learners(select = c("id", "mlr3_package", "required_package"))
```

# Installing and Loading Learners

The package includes functionality for detecting if you have the required packages installed
to use a learner, and ships with the function `install_learner` which can install all required
learner dependencies.

```{r, echo=FALSE, results='hide',error=TRUE,message=FALSE,warning=FALSE}
remove.packages("gbm")
unloadNamespace("gbm")
```
```{r,error=TRUE}
lrn("regr.gbm")
```
```{r, results='hide',error=TRUE,message=FALSE,warning=FALSE}
install_learners("regr.gbm")
```
```{r}
lrn("regr.gbm")
```

# Extending mlr3extralearners

New learners can be created with the `create_learner` function. This assumes you have a local
copy of `mlr3extralearners`. This function will automatically create the learner, learner tests,
parameter tests, YAML files for CI if required, and update the DESCRIPTION if required.
Once all tests are passing locally, open a [pull request](https://github.com/mlr-org/mlr3extralearners/pulls) with the "New Learner" template.

```{r, eval=FALSE}
create_learner(classname = "Locfit",
               algorithm = "localised fit",
               type = "dens",
               key = "locfit",
               package = "locfit",
               caller = "density.lf",
               feature_types = c("integer", "numeric"),
               predict_types = c("pdf", "cdf"),
               properties = NULL,
               importance = FALSE,
               oob_error = FALSE,
               references = FALSE)
```

