---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
library(mlr3extralearners)
library(magrittr)
```

# mlr3extralearners

Package website: [release](https://mlr3extralearners.mlr-org.com/) | [dev](https://mlr3extralearners.mlr-org.com/dev/)

Extra Learners for **[mlr3](https://github.com/mlr-org/mlr3/)**.

<!-- badges: start -->
[![R CMD Check](https://github.com/mlr-org/mlr3extralearners/workflows/R%20CMD%20Check/badge.svg)](https://mlr3extralearners.mlr-org.com/articles/learners/learner_status.html)
[![codecov](https://codecov.io/gh/mlr-org/mlr3extralearners/branch/main/graph/badge.svg)](https://codecov.io/gh/mlr-org/mlr3extralearners)
[![StackOverflow](https://img.shields.io/badge/stackoverflow-mlr3-orange.svg)](https://stackoverflow.com/questions/tagged/mlr3)
[![CodeFactor](https://www.codefactor.io/repository/github/mlr-org/mlr3extralearners/badge)](https://www.codefactor.io/repository/github/mlr-org/mlr3extralearners)
<!-- badges: end -->

## What is mlr3extralearners?

`mlr3extralearners` contains all learners from mlr3 that are not in `mlr3learners` or the
core packages. `mlr3extralearners` contains helper functions to find where all the learners, across the mlr3verse, live and to install required packages to run these learners. See the interactive [learner list](https://mlr3extralearners.mlr-org.com/articles/learners/list_learners.html) for the full list of learners in the mlr3verse and the [learner status page](https://mlr3extralearners.mlr-org.com/articles/learners/test_overview.html) for a live build status.

```{r}
list_mlr3learners(select = c("id", "mlr3_package", "required_packages"))
```

mlr3extralearners lives on GitHub and will not be on CRAN. Install with:

```{r, eval = FALSE}
remotes::install_github("mlr-org/mlr3extralearners")
```

## Installing and Loading Learners

The package includes functionality for detecting if you have the required packages installed
to use a learner, and ships with the function `install_learner` which can install all required
learner dependencies.

```{r, echo=FALSE, results='hide',error=TRUE,message=FALSE,warning=FALSE}
remove.packages("gbm")
unloadNamespace("gbm")
```
```{r,error=TRUE}
lrn("regr.gbm")
```
```{r, results='hide',error=TRUE,message=FALSE,warning=FALSE}
install_learners("regr.gbm")
```
```{r}
lrn("regr.gbm")
```

## Extending mlr3extralearners

New learners - either for personal use or to extend mlr3extralearners - can be created with the `create_learner` function.
An in-depth tutorial on how to do this can be found in the [mlr3 book](https://mlr3book.mlr-org.com/07-extending-learners.html).

